\setcounter{section}{0}

\vspace{20pt}

\begin{center}
    \textbf{Amendment I \\ MEMBERSHIP \& ELIGIBILITY \\ \textit{\footnotesize (approved by referendum on April 5, 2018)}}
\end{center}

\section{Rescission}

\textit{[Article IV, Section 6, Item (a); Article VII, Section 2, Item (b); and Article VII, Section 4, Item (a) are hereby rescinded]}

\section{Officer Eligibility}

The offices of Grand Marshal, President of the Union, and Chairman of the Judicial Board are open to any student who has been enrolled at Rensselaer for at least one full semester. The office of President of the Undergraduate Council is only available to an undergraduate student who has been enrolled at Rensselaer for at least one full semester. The office of President of the Graduate Council is only available to a graduate student who has been matriculated into a graduate program at Rensselaer for at least one full semester.

\vspace{12pt}

\textit{[This section shall be inserted as Article IV, Section 6, Item (a)]}

\section{Graduate Council}

The Graduate Council is the graduate legislative branch of the Union. Its voting membership shall consist of the six Graduate Senators and at least five representatives elected at large. It shall be presided over by the President of the Graduate Council.

\begin{enumerate}
    \item The number of graduate representatives, if greater than five, shall be defined in the bylaws of the Graduate Council.
    \item A vacancy caused by an increase in number of graduate representatives shall be filled by a person appointed by the Graduate Council.
    \item A reduction in number of graduate representatives shall only occur in the event of resignation, removal, or other vacancies allowing for the new number to occur.
    \item Quorum shall be adjusted to reflect a change in number of graduate representatives.
\end{enumerate}

\textit{[This section shall be inserted as Article VII, Section 4, Item (a)]}

\section{Executive Board Membership}

The Executive Board shall consist of a minimum of 15 voting members, and a maximum of 20 voting members:

\begin{enumerate}
    \item The following positions on the Executive Board must be filled: 5 club and organization representatives, 1 representative each from the Freshman, Sophomore, Junior, Senior, and Graduate classes, a representative of the Graduate Council, a representative of the Undergraduate Council, a Senate-Executive Board Liaison, and 2 members-at-large representing any membership of the Union.
    \item The following positions on the Executive Board may be filled at the discretion of the President of the Union: 2 club and organization members, and 3 members-at-large.
    \item The Senate-Executive Board Liaison, Graduate Council Representative, and Undergraduate Council Representative are appointed by the Senate, Graduate Council, and Undergraduate Council respectively. All other positions are appointed by the President of the Union and approved by a 2/3 vote of the Senate.
    \item If a member of the Executive Board is unable to fulfill his or her duties, a replacement may be appointed by the President of the Union, subject to a majority vote of the Executive Board and a 2/3 approval of the Senate.
    \item The Director of the Union and his or her assistants shall act as advisers to the Executive Board.
\end{enumerate}

\textit{[This section shall be inserted as Article VII, Section 2, Item (b)]}
