\article{OFFICERS}
\begin{enumerate}
\item The officers of the Union shall be the Grand Marshal, the President of the Union, the Chairperson of the Judicial Board, the President of the Graduate Council, and the President of the Undergraduate Council.
\item It shall be the duty of the Union officers to meet regularly to discuss policy and facilitate a closer coordination of their activities.
\end{enumerate}

\section{Grand Marshal}
The Grand Marshal shall be the presiding officer of the Student Senate and an ex-officio member of all committees.

\begin{enumerate}
\item The Grand Marshal shall be regarded as the leader and the chief spokesperson for the entire Union.
\item The Grand Marshal shall be a non-voting member of the Student Senate, but may cast one vote in the event of a tie.
\item Any appointments made by the Grand Marshal may be nullified by a majority vote of the total voting membership of the Student Senate.
\end{enumerate}

\section{President of the Union}
The President of the Union shall be the presiding officer of the Executive Board, and an ex-officio member of all Union committees.

\begin{enumerate}
\item The President of the Union shall be a non-voting member of the Executive Board but may cast one vote in the event of a tie.
\item Any appointments made by the President of the Union other than members of the Executive Board must be ratified by a $\frac{2}{3}$ vote of the Executive Board
\end{enumerate}

\section{Chairperson of the Judicial Board}
The Chairperson of the Judicial Board shall be the presiding officer of the Judicial Board.

\begin{enumerate}
\item The Chairperson shall be a non-voting member of the Judicial Board, but in the event of the tie may cast one vote.
\end{enumerate}

\section{President of the Graduate Council}
The President of the Graduate Council shall be the presiding officer of the Graduate Council and an ex-officio member of all graduate committees.

\begin{enumerate}
\item The President of the Graduate Council shall be elected by the Graduate Council from among its members.
\item His or her voting status will be determined via the by-laws of the Graduate Council.
\end{enumerate}

\section{President of the Undergraduate Council}
The President of the Undergraduate Council shall be the presiding officer of the Undergraduate Council and an ex-officio member of all undergraduate committees.

\begin{enumerate}
\item The President of the Undergraduate Council shall be a non-voting member of the
Undergraduate Council but may cast one vote in the case of a tie.
\end{enumerate}

\section{Term Limits}
\begin{enumerate}
\item The offices of Grand Marshal, President of the Union, and Chairman of the Judicial Board are open to any student who has been enrolled at Rensselaer for at least one full semester. The office of President of the Undergraduate Council is only available to an undergraduate student who has been enrolled at Rensselaer for at least one full semester. The office of President of the Graduate Council is only available to a graduate student who has been matriculated into a graduate program at Rensselaer for at least one full semester. \textit{[As amended by Amendment I.]}
\item A student may hold the position of either Grand Marshal or President of the Union for no more than a combined total of two full terms.
\end{enumerate}
