\article{REMOVALS}
\section{}
\begin{enumerate}
\item The Grand Marshal, the President of the Union, or the President of the Undergraduate Council
may be removed from office by a petition requiring his or her resignation signed by at least 50
percent of his or her constituents or by a $\frac{2}{3}$ vote of his or her constituents voting in a valid
removal election.
\item A valid removal election is one in which at least 20 percent of his or her constituents vote.
\item A removal election may be called by a $\frac{2}{3}$ vote of the Student Senate or by a petition signed by
at least 20 percent of his or her constituents followed by a majority vote of the Student Senate.
\end{enumerate}

\section{}
\begin{enumerate}
\item The Student Senate may remove a voting member for good cause, as specified in the Student
Senate by-laws, by a $\frac{2}{3}$ vote of its total voting membership.
\item Elected senators may be recalled by their constituencies. A $\frac{2}{3}$ vote of his or her constituents in
a valid removal election would constitute a recall. Such a removal election may be brought
about by a petition signed by 20 percent of his or her constituents or by a majority vote of the
Student Senate. Senators filling seats normally appointed by a council, may be removed by a
$\frac{2}{3}$ vote of the council.
\item Should a vacancy remain unfilled by the body which made the appointment that body shall
forfeit the vote of the unoccupied position.
\end{enumerate}

\section{}
\begin{enumerate}
\item Any members of the Executive Board whom have been installed by the Student Senate may be
removed by a $\frac{2}{3}$ vote of the Student Senate, or by the simple request of the President of the
Union supported by a majority vote of the Student Senate.
\item The members of the Executive Board appointed by the Graduate Council, Undergraduate
Council, or the Student Senate may be removed by a $\frac{2}{3}$ vote of the appointing body, or by the
simple request of the President of the Union supported by a majority vote of the appointing
body.
\end{enumerate}

\section{}
A student member of the Judicial Board or Review Board may be recalled for good cause by $\frac{2}{3}$ vote of the Student Senate.

\section{}
\begin{enumerate}
\item Any member of the Graduate Council may be removed from office for good cause as specified in
their by-laws, with a $\frac{2}{3}$ vote of the Graduate Council. Any member of the Undergraduate Class
Councils may be removed from office for good cause as specified in the Undergraduate Council
by-laws, by a $\frac{2}{3}$ of their total voting membership of their Class Council.
\item The person whom the action concerns shall not have the right to vote in the removal decision.
An impeachment vote, requiring majority approval, may be called from the floor at any Council
meeting. If the impeachment vote is successful, a removal vote will be held at the following
meeting. 
\end{enumerate}