\article{SUCCESSION}

\section{}
Should the position of Grand Marshal become vacant, within two weeks the Student Senate shall
appoint a new Grand Marshal from within its membership. His or her appointment shall require the
approval of $\frac{2}{3}$ vote of the Student Senate. The Chairperson of the Judicial Board will preside over this
appointment. Until the new Grand Marshal is appointed, the Student Senate by-laws shall dictate who
will assume the remaining duties of the Grand Marshal.

\section{}
Should the position of President of the Union become vacant, within two weeks the Executive Board
shall appoint from within its membership a new President. His or her appointment shall be confirmed
by a $\frac{2}{3}$ vote of the Student Senate. The Executive Board by-laws shall dictate who will preside over the
Executive Board until the new President of the Union is confirmed.

\section{}
Should the position of Judicial Board Chairman become vacant, within two weeks the Judicial Board shall
nominate a new Judicial Board Chairman from among the membership of the Judicial Board. The
nominee shall be confirmed by a two-thirds vote of the Senate. Until a new Chairman has been
confirmed, the Judicial Board by-laws shall dictate who will assume the duties of the Chairman.

\section{}
Should the position of the President of the Undergraduate Council become vacant, all active Class
Councils shall meet in joint session within two weeks, for the purpose of electing a new President of the
Undergraduate Council. His or her appointment shall require a $\frac{2}{3}$ approval of the Student Senate. The
Grand Marshal shall preside over this appointment. Until the new President of the Undergraduate
Council is appointed, the Undergraduate Council By-laws shall dictate who will assume the remaining
duties of the President.

\section{}
If any elective position including by not limited to, Officers of the Union, is filled by appointment, a
special election for that position shall be held if the Student Senate receives, within two weeks of the
announcement of the appointment, a petition signed by 20 percent of the constituency involved. If the
appointment is made at a time when the full student body is not normally on campus, the petition must
be received within two weeks of the announcement to the full campus.

\section{}
Any vacant Class Senator position shall be filled by a person appointed by their corresponding Class
Council.